% vim: set textwidth=78 autoindent:

\section{QGIS Coding Standards}

% when the revision of a section has been finalized,
% comment out the following line:
\updatedisclaimer

The following chapters provide coding information for QGIS Version \CURRENT. 
This document corresponds almost to a \LaTeX~ conversion of
the CODING.t2t file coming with the QGIS sources from December, 16th 2008.

These standards should be followed by all QGIS developers. Current information 
about QGIS Coding Standards are also available from wiki at:

\url{http://wiki.qgis.org/qgiswiki/CodingGuidelines} \\
\url{http://wiki.qgis.org/qgiswiki/CodingStandards} \\
\url{http://wiki.qgis.org/qgiswiki/UsingSubversion} \\
\url{http://wiki.qgis.org/qgiswiki/DebuggingPlugins} \\
\url{http://wiki.qgis.org/qgiswiki/DevelopmentInBranches} \\
\url{http://wiki.qgis.org/qgiswiki/SubmittingPatchesAndSvnAccess} \\

\subsection{Classes}
\subsubsection{Names}
Class in QGIS begin with Qgs and are formed using mixed case. 

\begin{verbatim}
Examples:
	QgsPoint
	QgsMapCanvas
	QgsRasterLayer
\end{verbatim}

\subsubsection{Members}
Class member names begin with a lower case \textit{m} and are formed using mixed case.

\begin{verbatim}
	mMapCanvas	
	mCurrentExtent
\end{verbatim}

All class members should be private.
\textbf{Public class members are STRONGLY discouraged}

\subsubsection{Accessor Functions}
Class member values should be obtained through accesssor functions. The function should be named without a \textit{get} prefix. Accessor functions for the two private members above would be: 

\begin{verbatim}
	mapCanvas()
	currentExtent()
\end{verbatim}

\subsubsection{Functions}
Function names begin with a lowercase letter and are formed using mixed case. The function name should convey something about the purpose of the function.

\begin{verbatim}
	updateMapExtent()
	setUserOptions()
\end{verbatim}

\subsection{Qt Designer}
\subsubsection{Generated Classes}
QGIS classes that are generated from Qt Designer (ui) files should have a \textit{Base} suffix. This identifies the class as a generated base class.

\begin{verbatim}
Examples:
	QgsPluginMangerBase
	QgsUserOptionsBase
\end{verbatim}
\subsubsection{Dialogs}
All dialogs should implement the following:
 * Tooltip help for all toolbar icons and other relevant widgets
 * WhatsThis help for \textbf{all} widgets on the dialog
 * An optional (though highly recommended) context sensitive \textit{Help} button that directs the user to the appropriate help page by launching their web browser

\subsection{C++ Files}
\subsubsection{Names}
C++ implementation and header files should be have a .cpp and .h extension respectively.
Filename should be all lowercase and, in the case of classes, match the class name.

\begin{verbatim}
Example:
	Class QgsFeatureAttribute source files are 
		qgsfeatureattribute.cpp and qgsfeatureattribute.h
\end{verbatim}

\subsubsection{Standard Header and License}
Each source file should contain a header section patterned after the following example:

\begin{verbatim}
/***************************************************************************
    qgsfield.cpp - Describes a field in a layer or table
     --------------------------------------
    Date                 : 01-Jan-2004
    Copyright            : (C) 2004 by Gary E.Sherman
    Email                : sherman at mrcc.com
/***************************************************************************
 *                                                                         *
 *   This program is free software; you can redistribute it and/or modify  *
 *   it under the terms of the GNU General Public License as published by  *
 *   the Free Software Foundation; either version 2 of the License, or     *
 *   (at your option) any later version.                                   *
 *                                                                         *
 ***************************************************************************/
\end{verbatim}

\subsubsection{CVS Keyword}
Each source file should contain the \$Id\$ keyword. This will be expanded by CVS to contain useful information about the file, revision, last committer, and date/time of last checkin.

Place the keyword right after the standard header/license that is found at the top of each source file:

\begin{verbatim}
	/* $Id$ */
\end{verbatim}

\subsection{Variable Names}
Variable names begin with a lower case letter and are formed using mixed case.

\begin{verbatim}
Examples:
	mapCanvas
	currentExtent
\end{verbatim}
\subsection{Enumerated Types}
Enumerated types should be named in CamelCase with a leading capital e.g.:

\begin{verbatim}
    enum UnitType
    {
      Meters,
      Feet,
      Degrees,
      UnknownUnit
    } ;
\end{verbatim}

Do not use generic type names that will conflict with other types. e.g. use "UnkownUnit" rather 
than "Unknown"

\subsection{Global Constants}
Global constants should be written in upper case underscore separated e.g.:

\begin{verbatim}
const long GEOCRS_ID = 3344;
\end{verbatim}

\subsection{Editing}
Any text editor/IDE can be used to edit QGIS code, providing the following requirements are met.

\subsubsection{Tabs}
Set your editor to emulate tabs with spaces. Tab spacing should be set to 2 spaces.

\subsubsection{Indentation}
Source code should be indented to improve readability. There is a .indent.pro file in the QGIS src directory that contains the switches to be used when indenting code using the GNU indent program. If you don't use GNU indent, you should emulate these settings.

\subsubsection{Braces}
Braces should start on the line following the expression:

\begin{verbatim}
	if(foo == 1)
	{
	  // do stuff
	  ...
 	}else
	{
	  // do something else
	  ...
	}
\end{verbatim}

\subsection{API Compatibility}
From QGIS 1.0 we will provide a stable, backwards compatible API. This will
provide a stable basis for people to develop against, knowing their code will
work against any of the 1.x QGIS releases (although recompiling may be
required).Cleanups to the API should be done in a manner similar to the
Trolltech developers e.g.

\begin{verbatim}
class Foo 
{
  public:
    /** This method will be deprecated, you are encouraged to use 
      doSomethingBetter() rather.
      @see doSomethingBetter()
     */
    bool doSomething();

    /** Does something a better way.
      @note This method was introduced in QGIS version 1.1
     */
    bool doSomethingBetter();

}
\end{verbatim}

\subsection{Coding Style}
Here are described some programming hints and tips that will hopefully reduce errors, development time, and maintenance.

\subsubsection{Where-ever Possible Generalize Code}

If you are cut-n-pasting code, or otherwise writing the same thing more than once, consider consolidating the code 
into a single function.

This will allow changes to be made in one location instead of in multiple places

\begin{itemize}
\item help prevent code bloat
\item make it more difficult for multiple copies to evolve differences over time, thus making it harder to understand and maintain for others
\end{itemize}

\subsubsection{Prefer Having Constants First in Predicates}
Prefer to put constants first in predicates. 

\begin{verbatim}
"0 == value" instead of "value == 0"
\end{verbatim}

This will help prevent programmers from accidentally using "=" when they meant to use "==", which can introduce very subtle 
logic bugs.  The compiler will generate an error if you accidentally use "=" instead of "==" for comparisons since constants 
inherently cannot be assigned values.

\subsubsection{Whitespace Can Be Your Friend}
Adding spaces between operators, statements, and functions makes it easier for humans to parse code.

Which is easier to read, this:

\begin{verbatim}
if (!a&&b)
\end{verbatim}

or this:

\begin{verbatim}
if ( ! a && b )
\end{verbatim}

\subsubsection{Add Trailing Identifying Comments}
Adding comments at the end of function, struct and class implementations makes it easier to find them later.

Consider that you're at the bottom of a source file and need to find a very long function -- without these kinds of trailing 
comments you will have to page up past the body of the function to find its name.  Of course this is ok if you wanted to find 
the beginning of the function; but what if you were interested at code near its end?  You'd have to page up and then back down 
again to the desired part.

E.g.,

\begin{verbatim}
void foo::bar()
{ 
    // ... imagine a lot of code here 
 } // foo::bar()
\end{verbatim}

\subsubsection{Use Braces Even for Single Line Statements}
Using braces for code in if/then blocks or similar code structures even for single line statements means that adding another 
statement is less likely to generate broken code.

Consider:

\begin{verbatim}
  if (foo)
    bar();
 else
    baz();
\end{verbatim}

Adding code after bar() or baz() without adding enclosing braces would create broken code.  Though most programmers would 
naturally do that, some may forget to do so in haste.

So, prefer this:

\begin{verbatim}
 if (foo)
 {
   bar();
 }
 else
 { 
    baz();
 } 
\end{verbatim}

\subsubsection{Book recommendations}

Effective C++ \url{http://www.awprofessional.com/title/0321334876} \\
More Effective C++ \url{http://www.awprofessional.com/bookstore/product.asp?isbn=020163371X&rl=1} \\
Effective STL \url{http://www.awprofessional.com/title/0201749629} \\
Design Patterns \url{http://www.awprofessional.com/title/0201634988}

You should also really read this article from Qt Quarterly on designing Qt style
\url{http://doc.trolltech.com/qq/qq13-apis.html}

\section{SVN Access}
This page describes how to get started using the QGIS Subversion repository

\subsection{Accessing the Repository}
To check out QGIS HEAD:

\begin{verbatim}
  svn --username [your user name] co https://svn.qgis.org/repos/qgis/trunk/qgis
\end{verbatim}

\subsection{Anonymous Access}
You can use the following commands to perform an anonymous checkout from the QGIS Subversion repository. 
Note we recommend checking out the trunk (unless you are a developer or really HAVE to have the latest 
changes and dont mind lots of crashing!).

You must have a subversion client installed prior to checking out the code. See the Subversion website 
for more information. The Links page contains a good selection of SVN clients for various platforms.

To check out a branch

\begin{verbatim}
  svn co https://svn.qgis.org/repos/qgis/branches/<branch name>
\end{verbatim}
To check out SVN stable trunk:

\begin{verbatim}
  svn co https://svn.qgis.org/repos/qgis/trunk/qgis qgis_unstable
\end{verbatim}

/!$\backslash$ \textbf{Note:} If you are behind a proxy server, edit your \~{}/subversion/servers file to specify 
your proxy settings first!

/!$\backslash$ \textbf{Note:} In QGIS we keep our most stable code in trunk. Periodically we will tag a release 
off trunk, and then continue stabilisation and selective incorporation of new features into trunk.

See the INSTALL file in the source tree for specific instructions on building development versions. 

\subsection{QGIS documentation sources}
If you're interested in checking out Quantum GIS documentation sources:

\begin{verbatim}
  svn co https://svn.qgis.org/repos/qgis_docs/trunk qgis_docs
\end{verbatim}

You can also take a look at DocumentationWritersCorner for more information.

\subsection{Documentation}
The repository is organized as follows:

\htmladdnormallink{http://wiki.qgis.org/images/repo.png}{http://wiki.qgis.org/images/repo.png}

See the Subversion book \htmladdnormallink{http://svnbook.red-bean.com}{http://svnbook.red-bean.com} for information on becoming a SVN master.

\subsection{Development in branches}
\subsubsection{Purpose}
The complexity of the QGIS source code has increased considerably during the last years. Therefore it is hard 
to anticipate the side effects that the addition of a feature will have. In the past, the QGIS project had very 
long release cycles because it was a lot of work to reetablish the stability of the software system after new 
features were added. To overcome these problems, QGIS switched to a development model where new features are 
coded in svn branches first and merged to trunk (the main branch) when they are finished and stable. This section 
describes the procedure for branching and merging in the QGIS project.

\subsubsection{Procedure}
 * Initial announcement on mailing list
Before starting, make an announcement on the developer mailing list to see if another developer is 
already working on the same feature. Also contact the technical advisor of the project steering committee 
(PSC). If the new feature requires any changes to the QGIS architecture, a request for comment (RFC) is needed. 
 * Create a branch
Create a new svn branch for the development of the new feature (see UsingSubversion for the svn syntax). Now 
you can start developing.
 * Merge from trunk regularly
It is recommended to merge the changes in trunk to the branch on a regular basis. This makes it easier to merge 
the branch back to trunk later.
 * Documentation on wiki
It is also recommended to document the intended changes and the current status of the work on a wiki page.
 * Testing before merging back to trunk
When you are finished with the new feature and happy with the stability, make an announcement on the developer list. 
Before merging back, the changes will be tested by developers and users. Binary packages (especially for OsX and Windows) 
will be generated to also involve non-developers. In trac, a new Component will be opened to file tickets against. 
Once there are no remaining issues left, the technical advisor of the PSC merges the changes into trunk.

\subsubsection{Creating a branch}
We prefer that new feature developments happen out of trunk so that trunk remains in a 
stable state. To create a branch use the following command:

\begin{verbatim}
svn copy https://svn.qgis.org/repos/qgis/trunk/qgis \
https://svn.qgis.org/repos/qgis/branches/qgis_newfeature
svn commit -m "New feature branch"
\end{verbatim}

\subsubsection{Merge regularly from trunk to branch}
When working in a branch you should regularly merge trunk into it so that your branch does not diverge more 
than necessary. In the top level dir of your branch, first type \texttt{`svn info`} to determine the revision 
numbers of your branch which will produce output something like this:

\begin{verbatim}
timlinux@timlinux-desktop:~/dev/cpp/qgis_raster_transparency_branch$ svn info
Caminho: .
URL: https://svn.qgis.org/repos/qgis/branches/raster_transparency_branch
Raiz do Repositorio: https://svn.qgis.org/repos/qgis
UUID do repositorio: c8812cc2-4d05-0410-92ff-de0c093fc19c
Revisao: 6546
Tipo de No: diretorio
Agendado: normal
Autor da Ultima Mudanca: timlinux
Revisao da Ultima Mudanca: 6495
Data da Ultima Mudanca: 2007-02-02 09:29:47 -0200 (Sex, 02 Fev 2007)
Propriedades da Ultima Mudanca: 2007-01-09 11:32:55 -0200 (Ter, 09 Jan 2007)
\end{verbatim}

The second revision number shows the revision number of the start revision of your branch and the first the 
current revision. You can do a dry run of the merge like this:

\begin{verbatim}
svn merge --dry-run -r 6495:6546 https://svn.qgis.org/repos/qgis/trunk/qgis
\end{verbatim}

After you are happy with the changes that will be made do the merge for real like this:

\begin{verbatim}
svn merge -r 6495:6546 https://svn.qgis.org/repos/qgis/trunk/qgis
svn commit -m "Merged upstream changes from trunk to my branch"
\end{verbatim}

\subsection{Submitting Patches}
There are a few guidelines that will help you to get your patches into QGIS easily, and help us 
deal with the patches that are sent to use easily.

\subsubsection{Patch file naming}
If the patch is a fix for a specific bug, please name the file with the bug number in it e.g. 
\textbf{bug777fix.diff}, and attach it to the original bug report in trac (\htmladdnormallink{https://trac.osgeo.org/qgis/}{https://trac.osgeo.org/qgis/}).

If the bug is an enhancement or new feature, its usually a good idea to create a ticket in 
trac (\htmladdnormallink{https://trac.osgeo.org/qgis/}{https://trac.osgeo.org/qgis/}) first and then attach you 

\subsubsection{Create your patch in the top level QGIS source dir}
This makes it easier for us to apply the patches since we don't need to navigate to a specific 
place in the source tree to apply the patch. Also when I receive patches I usually evaluate them 
using kompare, and having the patch from the top level dir makes this much easier. Below is an 
example of you you can include multiple changed files into your patch from the top level directory:

\begin{verbatim}
cd qgis
svn diff src/ui/somefile.ui src/app/somefile2.cpp > bug872fix.diff
\end{verbatim}

\subsubsection{Including non version controlled files in your patch}
If your improvements include new files that don't yet exist in the repository, you should indicate 
to svn that they need to be added before generating your patch e.g.

\begin{verbatim}
cd qgis
svn add src/lib/somenewfile.cpp
svn diff > bug7887fix.diff
\end{verbatim}

\subsubsection{Getting your patch noticed}
QGIS developers are busy folk. We do scan the incoming patches on bug reports but sometimes we miss things. 
Don't be offended or alarmed. Try to identify a developer to help you - using the ["Project Organigram"] and 
contact them asking them if they can look at your patch. If you dont get any response, you can escalate your 
query to one of the Project Steering Committee members (contact details also available on the ["Project Organigram"]).

\subsubsection{Due Diligence}
QGIS is licensed under the GPL. You should make every effort to ensure you only submit patches which are 
unencumbered by conflicting intellectual property rights. Also do not submit code that you are not happy to 
have made available under the GPL.

\subsection{Obtaining SVN Write Access}
Write access to QGIS source tree is by invitation. Typically when a person submits several (there is no fixed 
number here) substantial patches that demonstrate basic competance and understanding of C++ and QGIS coding 
conventions, one of the PSC members or other existing developers can nominate that person to the PSC for granting 
of write access. The nominator should give a basic promotional paragraph of why they think that person should gain 
write access. In some cases we will grant write access to non C++ developers e.g. for translators and documentors. 
In these cases, the person should still have demonstrated ability to submit patches and should ideally have submtted 
several substantial patches that demonstrate their understanding of modifying the code base without breaking things, etc.

\subsubsection{Procedure once you have access}
Checkout the sources:

\begin{verbatim}
svn co https://svn.qgis.org/repos/qgis/trunk/qgis qgis
\end{verbatim}

Build the sources (see INSTALL document for proper detailed instructions)

\begin{verbatim}
cd qgis
mkdir build
ccmake ..    (set your preferred options)
make
make install  (maybe you need to do with sudo / root perms)
\end{verbatim}

Make your edits

\begin{verbatim}
cd ..
\end{verbatim}

Make your changes in sources. Always check that everything compiles before making any commits.
Try to be aware of possible breakages your commits may cause for people building on other 
platforms and with older / newer versions of libraries.

Add files (if you added any new files). The svn status command can be used to quickly see 
if you have added new files.

\begin{verbatim}
svn status src/pluguns/grass/modules
\end{verbatim}

Files listed with ? in front are not in SVN and possibly need to be added by you:

\begin{verbatim}
svn add src/pluguns/grass/modules/foo.xml
\end{verbatim}

Commit your changes

\begin{verbatim}
svn commit src/pluguns/grass/modules/foo.xml
\end{verbatim}

Your editor (as defined in \$EDITOR environment variable) will appear and you should make a
 comment at the top of the file (above the area that says 'dont change this'. Put a 
descriptive comment and rather do several small commits if the changes across a number of 
files are unrelated. Conversely we prefer you to group related changes into a single commit.

Save and close in your editor. The first time you do this, you should be prompted to 
put in your username and password. Just use the same ones as your trac account.


\section{Unit Testing}
As of November 2007 we require all new features going into trunk to be accompanied with 
a unit test. Initially we have limited this requirement to qgis\_core, and we will extend 
this requirement to other parts of the code base once people are familiar with the 
procedures for unit testing explained in the sections that follow.

\subsection{The QGIS testing framework  - an overview}
Unit testing is carried out using a combination of QTestLib (the Qt testing library) and 
CTest (a framework for compiling and running tests as part of the CMake build process). 
Lets take an overview of the process before I delve into the details:

\begin{itemize}
\item \textbf{There is some code you want to test}, e.g. a class or function. Extreme programming 
   advocates suggest that the code should not even be written yet when you start 
   building your tests, and then as you implement your code you can immediately validate 
   each new functional part you add with your test. In practive you will probably 
   need to write tests for pre-existing code in QGIS since we are starting with a testing 
   framework well after much application logic has already been implemented.

\item \textbf{You create a unit test.} This happens under $<$QGIS Source Dir$>$/tests/src/core 
   in the case of the core lib. The test is basically a client that creates an instance 
   of a class and calls some methods on that class. It will check the return from each 
   method to make sure it matches the expected value. If any one of the calls fails,
   the unit will fail.

\item \textbf{You include QtTestLib macros in your test class.} This macro is processed by 
   the Qt meta object compiler (moc) and expands your test class into a runnable application. 

\item \textbf{You add a section to the CMakeLists.txt} in your tests directory that will
   build your test.

\item \textbf{You ensure you have ENABLE\_TESTING enabled in ccmake / cmakesetup.} This 
   will ensure your tests actually get compiled when you type make.

\item \textbf{You optionally add test data to $<$QGIS Source Dir$>$/tests/testdata} if your 
   test is data driven (e.g. needs to load a shapefile). These test data should be 
   as small as possible and wherever possible you should use the existing datasets 
   already there. Your tests should never modify this data in situ, but rather 
   may a temporary copy somewhere if needed.

\item \textbf{You compile your sources and install.} Do this using normal make \&\& (sudo) 
   make install procedure.

\item \textbf{You run your tests.} This is normally done simply by doing \textbf{make test} 
 after the make install step, though I will explain other aproaches that offer more 
 fine grained control over running tests.
\end{itemize}

Right with that overview in mind, I will delve into a bit of detail. I've already 
done much of the configuration for you in CMake and other places in the source tree 
so all you need to do are the easy bits - writing unit tests!

\subsection{Creating a unit test}
Creating a unit test is easy - typically you will do this by just creating a 
single .cpp file (not .h file is used) and implement all your test methods as
public methods that return void. I'll use a simple test class for QgsRasterLayer 
throughout the section that follows to illustrate. By convention we will name our 
test with the same name as the class they are testing but prefixed with 'Test'.
So our test implementation goes in a file called testqgsrasterlayer.cpp and 
the class itself will be TestQgsRasterLayer. First we add our standard copyright 
banner:

\newpage

\begin{verbatim}
/***************************************************************************
     testqgsvectorfilewriter.cpp
     --------------------------------------
    Date                 : Frida  Nov 23  2007
    Copyright            : (C) 2007 by Tim Sutton
    Email                : tim@linfiniti.com
 ***************************************************************************
 *                                                                         *
 *   This program is free software; you can redistribute it and/or modify  *
 *   it under the terms of the GNU General Public License as published by  *
 *   the Free Software Foundation; either version 2 of the License, or     *
 *   (at your option) any later version.                                   *
 *                                                                         *
 ***************************************************************************/
\end{verbatim}

Next we use start our includes needed for the tests we plan to run. There is 
one special include all tests should have:

\begin{verbatim}
#include <QtTest>
\end{verbatim}

Beyond that you just continue implementing your class as per normal, pulling 
in whatever headers you may need:

\begin{verbatim}
//Qt includes...
#include <QObject>
#include <QString>
#include <QObject>
#include <QApplication>
#include <QFileInfo>
#include <QDir>

//qgis includes...
#include <qgsrasterlayer.h> 
#include <qgsrasterbandstats.h> 
#include <qgsapplication.h>
\end{verbatim}

Since we are combining both class declaration and implementation in a single 
file the class declaration comes next. We start with our doxygen documentation. 
Every test case should be properly documented. We use the doxygen \textbf{ingroup} 
directive so that all the UnitTests appear as a module in the generated 
Doxygen documentation. After that comes a short description of the unit test:

\begin{verbatim}
/** \ingroup UnitTests
 * This is a unit test for the QgsRasterLayer class.
 */
\end{verbatim}

The class \textbf{must} inherit from QObject and include the Q\_OBJECT macro.

\begin{verbatim}
class TestQgsRasterLayer: public QObject
{
  Q_OBJECT;
\end{verbatim}

All our test methods are implemented as \textbf{private slots}. The QtTest framework 
will sequentially call each private slot method in the test class. There are 
four 'special' methods which if implemented will be called at the start of 
the unit test (\textbf{initTestCase}), at the end of the unit test (\textbf{cleanupTestCase}). 
Before each test method is called, the \textbf{init()} method will be called and 
after each test method is called the \textbf{cleanup()} method is called. These 
methods are handy in that they allow you to allocate and cleanup resources 
prior to running each test, and the test unit as a whole.

\begin{verbatim}
private slots:
    // will be called before the first testfunction is executed.
    void initTestCase();
    // will be called after the last testfunction was executed.
    void cleanupTestCase(){};
    // will be called before each testfunction is executed.
    void init(){};
    // will be called after every testfunction.
    void cleanup();
\end{verbatim}

Then come your test methods, all of which should take \textbf{no parameters} and 
should \textbf{return void}. The methods will be called in order of declaration.
I am implementing two methods here which illustrates to types of testing. In 
the first case I want to generally test the various parts of the class are 
working, I can use a \textbf{functional testing} approach. Once again, extreme 
programmers would advocate writing these tests \textbf{before} implementing the 
class. Then as you work your way through your class implementation you 
iteratively run your unit tests. More and more test functions should complete 
sucessfully as your class implementation work progresses, and when the whole 
unit test passes, your new class is done and is now complete with a repeatable 
way to validate it.

Typically your unit tests would only cover the \textbf{public} API of your 
class, and normally you do not need to write tests for accessors and mutators.
If it should happen that an acccessor or mutator is not working as expected 
you would normally implement a \textbf{regression} test to check for this (see 
lower down).

\begin{verbatim}
    //
    // Functional Testing
    //
    
    /** Check if a raster is valid. */
    void isValid();

    // more functional tests here ...
\end{verbatim}

Next we implement our \textbf{regression tests}. Regression tests should be 
implemented to replicate the conditions of a particular bug. For example 
I recently received a report by email that the cell count by rasters was 
off by 1, throwing off all the statistics for the raster bands. I opened 
a bug (ticket \#832) and then created a regression test that replicated 
the bug using a small test dataset (a 10x10 raster). Then I ran the test 
and ran it, verifying that it did indeed fail (the cell count was 99 
instead of 100). Then I went to fix the bug and reran the unit test and 
the regression test passed. I committed the regression test along with 
the bug fix. Now if anybody breakes this in the source code again in the 
future, we can immediatly identify that the code has regressed. Better 
yet before committing any changes in the future, running our tests will 
ensure our changes dont have unexpected side effects - like breaking 
existing functionality.

There is one more benifit to regression tests - they can save you time. 
If you ever fixed a bug that involved making changes to the source, 
and then running the application and performing a series of convoluted 
steps to replicate the issue, it will be immediately apparent that 
simply implementing your regression test \textbf{before} fixing the bug 
will let you automate the testing for bug resolution in an efficient 
manner.

To implement your regression test, you should follow the naming 
convention of regression$<$TicketID$>$ for your test functions. If no 
trac ticket exists for the regression, you should create one first. 
Using this approach allows the person running a failed regression 
test easily go and find out more information.

\begin{verbatim}
    //
    // Regression Testing
    //
    
    /** This is our second test case...to check if a raster
     reports its dimensions properly. It is a regression test 
     for ticket #832 which was fixed with change r7650. 
     */
    void regression832(); 
  
    // more regression tests go here ...
\end{verbatim}

Finally in our test class declaration you can declare privately 
any data members and helper methods your unit test may need. In our 
case I will declare a QgsRasterLayer * which can be used by any 
of our test methods. The raster layer will be created in the 
initTestCase() function which is run before any other tests, and then
destroyed using cleanupTestCase() which is run after all tests. By 
declaring helper methods (which may be called by various test 
functions) privately, you can ensure that they wont be automatically 
run by the QTest executeable that is created when we compile our test.

\begin{verbatim}
  private:
    // Here we have any data structures that may need to 
    // be used in many test cases.
    QgsRasterLayer * mpLayer;
};

\end{verbatim}

That ends our class declaration. The implementation is simply 
inlined in the same file lower down. First our init and cleanup functions:

\begin{verbatim}
void TestQgsRasterLayer::initTestCase()
{
  // init QGIS's paths - true means that all path will be inited from prefix
  QString qgisPath = QCoreApplication::applicationDirPath ();
  QgsApplication::setPrefixPath(qgisPath, TRUE);
#ifdef Q_OS_LINUX
  QgsApplication::setPkgDataPath(qgisPath + "/../share/qgis");
#endif
  //create some objects that will be used in all tests...

  std::cout << "Prefix  PATH: " << QgsApplication::prefixPath().toLocal8Bit().data() \
  << std::endl;
  std::cout << "Plugin  PATH: " << QgsApplication::pluginPath().toLocal8Bit().data() \
  << std::endl;
  std::cout << "PkgData PATH: " << QgsApplication::pkgDataPath().toLocal8Bit().data() \
  << std::endl;
  std::cout << "User DB PATH: " << QgsApplication::qgisUserDbFilePath().toLocal8Bit() \
  .data() << std::endl;

  //create a raster layer that will be used in all tests...
  QString myFileName (TEST_DATA_DIR); //defined in CmakeLists.txt
  myFileName = myFileName + QDir::separator() + "tenbytenraster.asc";
  QFileInfo myRasterFileInfo ( myFileName );
  mpLayer = new QgsRasterLayer ( myRasterFileInfo.filePath(),
            myRasterFileInfo.completeBaseName() );
}

void TestQgsRasterLayer::cleanupTestCase()
{
  delete mpLayer;
}

\end{verbatim}

The above init function illustrates a couple of interesting things.

 1. I needed to manually set the QGIS application data path so that
   resources such as srs.db can be found properly.
 2. Secondly, this is a data driven test so we needed to provide a 
   way to generically locate the 'tenbytenraster.asc file. This was 
   achieved by using the compiler define \textbf{TEST\_DATA\_PATH}. The 
   define is created in the CMakeLists.txt configuration file under 
   $<$QGIS Source Root$>$/tests/CMakeLists.txt and is available to all 
   QGIS unit tests. If you need test data for your test, commit it 
   under $<$QGIS Source Root$>$/tests/testdata. You should only commit 
   very small datasets here. If your test needs to modify the test 
   data, it should make a copy of if first.

Qt also provides some other interesting mechanisms for data driven 
testing, so if you are interested to know more on the topic, consult 
the Qt documentation.

Next lets look at our functional test. The isValid() test simply 
checks the raster layer was correctly loaded in the initTestCase. 
QVERIFY is a Qt macro that you can use to evaluate a test condition. 
There are a few other use macros Qt provide for use in your tests 
including:

\begin{verbatim}
QCOMPARE ( actual, expected )
QEXPECT_FAIL ( dataIndex, comment, mode )
QFAIL ( message )
QFETCH ( type, name )
QSKIP ( description, mode )
QTEST ( actual, testElement )
QTEST_APPLESS_MAIN ( TestClass )
QTEST_MAIN ( TestClass )
QTEST_NOOP_MAIN ()
QVERIFY2 ( condition, message )
QVERIFY ( condition )
QWARN ( message ) 
\end{verbatim}

Some of these macros are useful only when using the Qt framework 
for data driven testing (see the Qt docs for more detail).

\begin{verbatim}
void TestQgsRasterLayer::isValid()
{
  QVERIFY ( mpLayer->isValid() );
}
\end{verbatim}

Normally your functional tests would cover all the range of 
functionality of your classes public API where feasible. With our 
functional tests out the way, we can look at our regression test example.

Since the issue in bug \#832 is a misreported cell count, writing 
our test if simply a matter of using QVERIFY to check that the 
cell count meets the expected value:

\begin{verbatim}
void TestQgsRasterLayer::regression832()
{
   QVERIFY ( mpLayer->getRasterXDim() == 10 );
   QVERIFY ( mpLayer->getRasterYDim() == 10 );
   // regression check for ticket #832
   // note getRasterBandStats call is base 1
   QVERIFY ( mpLayer->getRasterBandStats(1).elementCountInt == 100 );
}
\end{verbatim}

With all the unit test functions implemented, there one final thing we 
need to add to our test class:

\begin{verbatim}
QTEST_MAIN(TestQgsRasterLayer)
#include "moc_testqgsrasterlayer.cxx"
\end{verbatim}

The purpose of these two lines is to signal to Qt's moc that his is a 
QtTest (it will generate a main method that in turn calls each test funtion.
The last line is the include for the MOC generated sources. You should 
replace 'testqgsrasterlayer' with the name of your class in lower case.

\subsection{Adding your unit test to CMakeLists.txt}
Adding your unit test to the build system is simply a matter of editing 
the CMakeLists.txt in the test directory, cloning one of the existing 
test blocks, and then search and replacing your test class name into it. 
For example:

\begin{verbatim}
#
# QgsRasterLayer test
#
SET(qgis_rasterlayertest_SRCS testqgsrasterlayer.cpp)
SET(qgis_rasterlayertest_MOC_CPPS testqgsrasterlayer.cpp)
QT4_WRAP_CPP(qgis_rasterlayertest_MOC_SRCS ${qgis_rasterlayertest_MOC_CPPS})
ADD_CUSTOM_TARGET(qgis_rasterlayertestmoc ALL DEPENDS ${qgis_rasterlayertest_MOC_SRCS})
ADD_EXECUTABLE(qgis_rasterlayertest ${qgis_rasterlayertest_SRCS})
ADD_DEPENDENCIES(qgis_rasterlayertest qgis_rasterlayertestmoc)
TARGET_LINK_LIBRARIES(qgis_rasterlayertest ${QT_LIBRARIES} qgis_core)
INSTALL(TARGETS qgis_rasterlayertest RUNTIME DESTINATION ${QGIS_BIN_DIR})
ADD_TEST(qgis_rasterlayertest ${QGIS_BIN_DIR}/qgis_rasterlayertest)
\end{verbatim}

I'll run through these lines briefly to explain what they do, but if 
you are not interested, just clone the block, search and replace e.g.

\begin{verbatim}
:'<,'>s/rasterlayer/mynewtest/g
\end{verbatim}

Lets look a little more in detail at the individual lines. First we 
define the list of sources for our test. Since we have only one source file 
(following the methodology I described above where class declaration and 
definition are in the same file) its a simple statement:

\begin{verbatim}
SET(qgis_rasterlayertest_SRCS testqgsrasterlayer.cpp)
\end{verbatim}

Since our test class needs to be run through the Qt meta object compiler (moc) 
we need to provide a couple of lines to make that happen too:

\begin{verbatim}
SET(qgis_rasterlayertest_MOC_CPPS testqgsrasterlayer.cpp)
QT4_WRAP_CPP(qgis_rasterlayertest_MOC_SRCS ${qgis_rasterlayertest_MOC_CPPS})
ADD_CUSTOM_TARGET(qgis_rasterlayertestmoc ALL DEPENDS ${qgis_rasterlayertest_MOC_SRCS})
\end{verbatim}

Next we tell cmake that it must make an executeable from the test class. 
Remember in the previous section on the last line of the class implementation 
I included the moc outputs directly into our test class, so that will 
give it (among other things) a main method so the class can be 
compiled as an executeable:

\begin{verbatim}
ADD_EXECUTABLE(qgis_rasterlayertest ${qgis_rasterlayertest_SRCS})
ADD_DEPENDENCIES(qgis_rasterlayertest qgis_rasterlayertestmoc)
\end{verbatim}

Next we need to specify any library dependencies. At the moment classes 
have been implemented with a catch-all QT\_LIBRARIES dependency, but I will 
be working to replace that with the specific Qt libraries that each class 
needs only. Of course you also need to link to the relevant qgis 
libraries as required by your unit test.

\begin{verbatim}
TARGET_LINK_LIBRARIES(qgis_rasterlayertest ${QT_LIBRARIES} qgis_core)
\end{verbatim}

Next I tell cmake to the same place as the qgis binaries itself. This 
is something I plan to remove in the future so that the tests can 
run directly from inside the source tree.

\begin{verbatim}
INSTALL(TARGETS qgis_rasterlayertest RUNTIME DESTINATION ${QGIS_BIN_DIR})
\end{verbatim}

Finally here is where the best magic happens - we register the class with 
ctest. If you recall in the overview I gave in the beginning of this 
section we are using both QtTest and CTest together. To recap, \textbf{QtTest} adds a 
main method to your test unit and handles calling your test methods within 
the class. It also provides some macros like QVERIFY that you can use as 
to test for failure of the tests using conditions. The output from 
a QtTest unit test is an executeable which you can run from the command line. 
However when you have a suite of tests and you want to run each executeable 
in turn, and better yet integrate running tests into the build process, 
the \textbf{CTest} is what we use. The next line registers the unit test with 
CMake / CTest.

\begin{verbatim}
ADD_TEST(qgis_rasterlayertest ${QGIS_BIN_DIR}/qgis_rasterlayertest)
\end{verbatim}

The last thing I should add is that if your test requires optional 
parts of the build process (e.g. Postgresql support, GSL libs, GRASS etc.), 
you should take care to enclose you test block inside a IF () block 
in the CMakeLists.txt file.

\subsection{Building your unit test}
To build the unit test you need only to make sure that ENABLE\_TESTS=true 
in the cmake configuration. There are two ways to do this:

 1. Run ccmake .. (cmakesetup .. under windows) and interactively set 
 the ENABLE\_TESTS flag to ON.
 1. Add a command line flag to cmake e.g. cmake -DENABLE\_TESTS=true ..

Other than that, just build QGIS as per normal and the tests should build 
too.

\subsection{Run your tests}
The simplest way to run the tests is as part of your normal build process:

\begin{verbatim}
make && make install && make test
\end{verbatim}

The make test command will invoke CTest which will run each test that 
was registered using the ADD\_TEST CMake directive described above. Typical 
output from make test will look like this:

\begin{verbatim}
Running tests...
Start processing tests
Test project /Users/tim/dev/cpp/qgis/build
1/  3 Testing qgis_applicationtest          ***Exception: Other
2/  3 Testing qgis_filewritertest           *** Passed
3/  3 Testing qgis_rasterlayertest          *** Passed

0% tests passed, 3 tests failed out of 3

  The following tests FAILED:
  1 - qgis_applicationtest (OTHER_FAULT)
  Errors while running CTest
  make: *** [test] Error 8
\end{verbatim}

If a test fails, you can use the ctest command to examine more 
closely why it failed. User the -R option to specify a regex for 
which tests you want to run and -V to get verbose output:

\begin{verbatim}
[build] ctest -R appl -V
Start processing tests
Test project /Users/tim/dev/cpp/qgis/build
Constructing a list of tests
Done constructing a list of tests
Changing directory into /Users/tim/dev/cpp/qgis/build/tests/src/core
1/  3 Testing qgis_applicationtest          
Test command: /Users/tim/dev/cpp/qgis/build/tests/src/core/qgis_applicationtest
********* Start testing of TestQgsApplication *********
  Config: Using QTest library 4.3.0, Qt 4.3.0
PASS   : TestQgsApplication::initTestCase()
  Prefix  PATH: /Users/tim/dev/cpp/qgis/build/tests/src/core/../
  Plugin  PATH: /Users/tim/dev/cpp/qgis/build/tests/src/core/..//lib/qgis
  PkgData PATH: /Users/tim/dev/cpp/qgis/build/tests/src/core/..//share/qgis
  User DB PATH: /Users/tim/.qgis/qgis.db
PASS   : TestQgsApplication::getPaths()
  Prefix  PATH: /Users/tim/dev/cpp/qgis/build/tests/src/core/../
  Plugin  PATH: /Users/tim/dev/cpp/qgis/build/tests/src/core/..//lib/qgis
  PkgData PATH: /Users/tim/dev/cpp/qgis/build/tests/src/core/..//share/qgis
  User DB PATH: /Users/tim/.qgis/qgis.db
  QDEBUG : TestQgsApplication::checkTheme() Checking if a theme icon exists:
  QDEBUG : TestQgsApplication::checkTheme() 
  /Users/tim/dev/cpp/qgis/build/tests/src/core/.. \
  //share/qgis/themes/default//mIconProjectionDisabled.png
  FAIL!  : TestQgsApplication::checkTheme() '!myPixmap.isNull()' returned FALSE. ()
  Loc: [/Users/tim/dev/cpp/qgis/tests/src/core/testqgsapplication.cpp(59)]
PASS   : TestQgsApplication::cleanupTestCase()
  Totals: 3 passed, 1 failed, 0 skipped
  ********* Finished testing of TestQgsApplication *********
  -- Process completed
  ***Failed

  0% tests passed, 1 tests failed out of 1

  The following tests FAILED:
1 - qgis_applicationtest (Failed)
  Errors while running CTest

\end{verbatim}

Well that concludes this section on writing unit tests in QGIS. We hope you 
will get into the habit of writing test to test new functionality and to 
check for regressions. Some aspects of the test system (in particular the 
CMakeLists.txt parts) are still being worked on so that the testing framework 
works in a truly platform way. I will update this document as things progress.


\section{HIG (Human Interface Guidelines)}
In order for all graphical user interface elements to appear consistant and 
to all the user to instinctively use dialogs, it is important that the following 
guidelines are followed in layout and design of GUIs.

 \begin{enumerate}
 \item Group related elements using group boxes:
   Try to identify elements that can be grouped together and then use 
   group boxes with a label to identify the topic of that group. 
   Avoid using group boxes with only a single widget / item inside.
 \item Capitalise first letter only in labels:
   Labels (and group box labels) should be written as a phrase with leading capital letter,
   and all remaing words written with lower case first letters 
 \item Do not end labels for widgets or group boxes with a colon:
   Adding a colon causes visual noise and does not impart additional meaning,
   so dont use them. An exception to this rule is when you have two labels 
   next to each other e.g.: Label1 \htmladdnormallink{Plugin}{Path:} Label2 [/path/to/plugins]
 \item Keep harmful actions away from harmless ones:
   If you have actions for 'delete', 'remove' etc, try to impose adequate 
   space between the harmful action and innocuous actions so that the users 
   is less likely to inadvertantly click on the harmful action.
 \item Always use a QButtonBox for 'OK', 'Cancel' etc buttons:
   Using a button box will ensure that the order of 'OK' and 'Cancel' etc, 
   buttons is consistent with the operating system / locale / desktop 
   environment that the user is using.
 \end{enumerate}


