% vim: set textwidth=78 autoindent:

\section{Forward}\label{label_forward}
\pagenumbering{arabic}
\setcounter{page}{1}

% when the revision of a section has been finalized, 
% comment out the following line:
\updatedisclaimer

Welcome to the wonderful world of Geographical Information Systems (GIS)!
Quantum GIS (QGIS) is an Open Source Geographic Information System. The project
was born in May of 2002 and was established as a project on SourceForge in June
of the same year. We've worked hard to make GIS software (which is traditionaly
expensive proprietary software) a viable prospect for anyone with basic access
to a Personal Computer. QGIS currently runs on most Unix platforms, Windows, and
OS X. QGIS is developed using the Qt toolkit (\url{http://www.trolltech.com})
and C++. This means that QGIS feels snappy to use and has a pleasing, easy to
use graphical user interface. 

QGIS aims to be an easy-to-use GIS, providing common functions and features.
The initial goal was to provide a GIS data viewer. QGIS has reached that point
in its evolution and is being used by many for their daily GIS data viewing
needs. QGIS supports a number of raster and vector data formats, with new
support easily added using the plugin architecture (see Appendix
\ref{appdx_data_formats} for a full list of currently supported data formats).
QGIS is released under the GNU General Public License (GPL). Developing QGIS 
under this license means that you can inspect and modify the source code,
and guarantees that you, our happy user, will always have access to a GIS
program that is free of cost and can be freely modified. You should have
received a full copy of the license with your copy of QGIS, and you also can
find it as Appendix \ref{gpl_appendix}.  

\begin{quote}
\begin{center}
\textbf{Note:} The latest version of this document can always be found at \newline
\url{http://download.osgeo.org/qgis/doc/manual/}
\end{center}
\end{quote}

\subsection{Features}\label{label_majfeat}

QGIS offers many common GIS functionalities provided by core features and
plugins. As a short summary they are presented in six categories to gain a
first insight.

\minisec{View data}

You can view and overlay vector and raster data in different formats and
projections without conversion to an internal or common format. Supported
formats include:

\begin{itemize}
\item spatially-enabled PostgreSQL tables using PostGIS, \textbf{most} vector
formats\footnote{OGR-supported database formats are not yet supported in QGIS.}
supported by the OGR library, including ESRI shapefiles, MapInfo,
SDTS and GML.
\item all raster formats supported by the GDAL library, such as digital
elevation models, aerial photography or landsat imagery,
\item GRASS mapsets, 
\item Online spatial data served as OGC-compliant WMS or WFS.
\end{itemize}

\minisec{Explore data and compose maps} 

You can compose maps and interactively explore spatial data with a friendly
graphical user interface. The many helpful tools available in the GUI include:

\begin{itemize}
\item on the fly projection
\item map composer
\item overview panel
\item spatial bookmarks
\item identify/select features
\item edit/view/search attributes
\item feature labeling
\item change vector and raster symbology
\item add a graticule layer
\item decorate your map with a north arrow scale bar and copyright label
\item save and restore projects
\end{itemize}

\minisec{Create, edit and export data}

You can create, edit and export vector maps in several formats. Raster data
have to imported into GRASS to be able to edit and export them into other
formats.  

\begin{itemize}
\item digitizing tools for OGR formats and GRASS vector
\item creating and editing shapefiles
\item geocode images with the georeferencer plugin
\item using the GPS tools to import and export GPX format, convert other GPS
formats to GPX or down/upload directly to a GPS unit
\item creating PostGIS layers from shapefiles with the SPIT plugin 
\end{itemize}

\minisec{Analyse data} 

You can perform spatial data analysis on PostgreSQL/PostGIS and other OGR
supported formats using the geoprocessing plugin. It currently offers buffer,
intersection, union, dissolve, difference and convex hull. Or you use the 
integrated GRASS tools, including the complete GRASS functionality, e.g:

\begin{itemize}
\item Image processing (image fusion, classification, filtering, enhancement)
\item Raster analysis (map algebra, terrain analysis, hydrologic modeling
\item Vector analysis (digitizing, union, intersection, buffer)
\item Network analysis (shortest path, travelling salesman)
\end{itemize}

\minisec{Publish maps on the internet}

QGIS can be used to export data to a mapfile and to visualize them in the
internet using a webserver with UMN MapServer installed. QGIS can also
be used as a WMS and WFS client and WMS server. 

\minisec{Extend QGIS functionality through plugins} 

Quantum GIS can be adapted to your special needs with the extensible
plugin architecture, providing libraries that can be used to create plugins
or even new applications with C++ or Python.     

\begin{itemize}
\item \textbf{Core Plugins}
\\ \\ Add WFS Layer
\\ Add Delimited Text Layer
\\ Decorations (Copyright Label, North Arrow and Scale bar)
\\ Georeferencer
\\ GPS Tools
\\ GRASS integration
\\ Graticules Creator
\\ SPIT Shapefile to PostgreSQL/PostGIS Import Tool
\\ Python Console
\\ Python Plugin Installer
\\ \item \textbf{Python Plugins}
\\ \\ QGIS offers a growing number of python plugins, provided by the
community. These plugins can be browsed and installed from the official
PyQGIS repository using the python plugin installer.
\end{itemize}

%% FIXME (maybe the whats new part should be added again?)






