% vim: set textwidth=78 autoindent:

% QGIS Tips
% define tip float
% doesn't work if written in qgis_style.sty
% please keep the style definitions here and 
% and load float package in qgis_style.sty
\floatstyle{ruled}
\newfloat{Tip}{ht}{lox}
\floatname{Tip}{Tip}
\newcommand\qgistip[1]{\raggedright\small{#1}}
\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}

\thispagestyle{empty}
\addcontentsline{toc}{section}{Preamble}

%%%%%%%%%%% nothing to change above %%%%%%%%%%

\section*{Preamble}

% when the revision of a section has been finalized, 
% comment out the following line:
%\updatedisclaimer

This document is the original user guide of the described 
software Quantum GIS. The software and hardware described in this 
document are in most cases registered trademarks and are therefore subject 
to the legal requirements. Quantum GIS is subject to the GNU General Public 
License. Find more information on the Quantum GIS Homepage
\url{http://qgis.osgeo.org}.

The details, data, results etc.~in this document have been 
written and verified to the best of knowledge and responsibility of the 
authors and editors. Nevertheless, mistakes concerning the content are possible.

Therefore, all data are not liable to any duties or guarantees. The authors, editors 
and publishers do not take any responsibility or liability for failures and 
their consequences. Your are always welcome to indicate possible mistakes.

This document has been typeset with \LaTeX. It is available as \LaTeX~source
code via \href{http://wiki.qgis.org/qgiswiki/DocumentationWritersCorner}{subversion} 
and online as PDF document via \url{http://qgis.osgeo.org/documentation/manuals.html}. 
Translated versions of this document can be downloaded via the documentation 
area of the QGIS project as well. For more information about contributing to
this document and about translating it, please visit: \url{http://wiki.qgis.org/qgiswiki/DocumentationWritersCorner} 

\vspace{0.5cm}

\textbf{Links in this Document}

This document contains internal and external links.  Clicking on an
internal link moves within the document, while clicking on an external link
opens an internet address.  In PDF form, internal links are shown in blue,
while external links are shown in red and are handled by the
system browser. In HTML form, the browser displays and handles both
identically. 

\begin{flushleft}
\textbf{User, Installation and Coding Guide Authors and Editors:}
 
\begin{tabular}{p{4cm} p{4cm} p{4cm} p{4cm}}
Tara Athan & Radim Blazek & Godofredo Contreras & Claudia A. Engel \\
Otto Dassau & Martin Dobias & J\"urgen E. Fischer & Anne Ghisla \\ 
Stephan Holl & Marco Hugentobler & Magnus Homann & Tim Sutton \\ 
Lars Luthman & Gavin Macaulay & Werner Macho & Carson J.Q. Farmer \\
Tyler Mitchell & Brendan Morely & Gary E. Sherman & David Willis \\ 
\end{tabular}

\vspace{1cm}

With thanks to Tisham Dhar for preparing the initial msys (MS Windows)
environment documentation, to Tom Elwertowski and William Kyngesburye for
help in the MAC OSX Installation Section and to Carlos D\'{a}vila, Paolo
Cavallini and Christian Gunning for revisions. If we have neglected to 
mention any contributors, please accept our apologies for this oversight.

\textbf{Copyright \copyright~2004 - 2009 Quantum GIS Development Team} \\
\textbf{Internet:} \url{http://qgis.osgeo.org}
\end{flushleft}

